---
title: "design_experiments"
output: html_document
date: '2026-04-26'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Design of experiments

Designing an experiment carefully is just as important as the analysis that follows it. R has a robust ecosystem of packages dedicated to Design of Experiments (DoE), catering to everything from agricultural field trials to industrial quality control and algorithmic optimization.

Here is a breakdown of the top R packages for designing experiments and generating layouts, complete with worked examples.

## 1. agricolae: Best for Classical & Field Designs
Originally designed for agricultural research, agricolae is an excellent package for classical experimental designs like Randomized Complete Block Designs (RCBD), Latin Squares, and Split-Plots. It not only generates the randomized design but also provides spatial "sketches" of the layout, which is incredibly useful for physical experiments.

Worked Example: Randomized Complete Block Design (RCBD)
Imagine you are testing four different fertilizers (A, B, C, D) across 4 different blocks (fields) to account for soil gradient differences.

```{r agricolae, echo=FALSE}
# Install if necessary: install.packages("agricolae")
library(agricolae)

# Define treatments and number of blocks (replications)
treatments <- c("Fertilizer_A", "Fertilizer_B", "Fertilizer_C", "Fertilizer_D")
blocks <- 4

# Generate the RCBD
# seed is set for reproducibility
my_design <- design.rcbd(trt = treatments, r = blocks, seed = 42)

# 1. View the 'book' (The actual randomized data frame for data collection)
print("--- Data Collection Book ---")
head(my_design$book)

# 2. View the 'sketch' (The physical layout of the experiment)
print("--- Physical Field Layout ---")
print(my_design$sketch)
```

## 2. FrF2 & DoE.base

 Best for Industrial & Factorial DesignsWhen you have multiple factors (variables) and need to see how they interact without running thousands of trials, you use Fractional Factorial designs. The FrF2 package specializes in 2-level fractional factorial designs, which are the backbone of industrial screening experiments, Six Sigma, and A/B/n testing.Worked Example: A $2^{5-1}$ Fractional Factorial DesignSuppose you are optimizing a chemical reaction and have 5 factors (Temperature, Pressure, Time, Catalyst, Stir Rate). A full factorial would require $2^5 = 32$ runs. If your budget only allows for 16 runs, you can create a "half-fraction" design.

```{r doe, echo=FALSE}
# Install if necessary: install.packages("FrF2")
library(FrF2)

# Generate a 16-run design for 5 factors
# We can name the factors and set their high/low levels
chem_design <- FrF2(nruns = 16, 
                    nfactors = 5, 
                    factor.names = list(
                      Temp = c(100, 200),     # 100°C to 200°C
                      Pressure = c(1, 5),     # 1 atm to 5 atm
                      Time = c(30, 60),       # 30 mins to 60 mins
                      Catalyst = c("A", "B"), # Type A or B
                      StirRate = c(100, 500)  # 100 rpm to 500 rpm
                    ),
                    seed = 123)

# View the randomized run order and factor settings
print(chem_design)

# View the design summary (shows confounding/aliasing structure)
summary(chem_design)
```

Why this is useful: FrF2 automatically handles the complex math of "aliasing" (ensuring your main effects are not confused with your two-way interactions) and gives you a ready-to-execute list of runs.

#3. algDesign:

Best for Custom & Algorithmic DesignsSometimes standard designs (like RCBD or simple factorials) don't fit. Maybe you have a weird sample size, constraints on your combinations (e.g., you can't have High Temperature AND High Pressure), or categorical variables with 3 levels mixed with continuous variables. algDesign uses computer algorithms to find the "optimal" design (usually D-optimal) from a candidate set of all possible runs.Worked Example: D-Optimal Design for a Constrained BudgetYou have 3 factors, each with 3 levels. A full factorial is $3 \times 3 \times 3 = 27$ runs. You only have the budget to run exactly 12 experiments, and you want to estimate a quadratic model.

```{r alpha, echo=FALSE}
# Install if necessary: install.packages("algDesign")
library(algDesign)

# Step 1: Generate the full candidate space (all 27 possible combinations)
candidate_set <- gen.factorial(levels = 3, nVars = 3, varNames = c("X1", "X2", "X3"))

# Step 2: Use the Federov exchange algorithm to find the best 12 runs
# We specify the model we want to fit: ~ quad(X1, X2, X3) means we want main effects and quadratic terms
optimal_design <- optFederov(~ quad(X1, X2, X3), 
                             data = candidate_set, 
                             nTrials = 12, 
                             evaluateI = TRUE)

# View the selected 12 runs
print(optimal_design$design)
```

| Package | Best Use Case | Common Designs |
| :--- | :--- | :--- |
| **`agricolae`** | Agriculture, ecology, biology where physical layout matters. | RCBD, Latin Square, Split-plot, Alpha designs. |
| **`FrF2` & `DoE.base`** | Engineering, manufacturing, software A/B testing (screening many factors). | 2-level full and fractional factorials, Plackett-Burman. |
| **`algDesign`** | Experiments with tight budgets, weird sample sizes, or constraints. | D-optimal, A-optimal, I-optimal designs. |
| **`edibble`** | Modern R users wanting to integrate DoE cleanly into the `tidyverse`. | Grammar of experimental design (tidy-compliant layouts). |

