---
title: "Diallel Analysis Using R"
output: html_document
date: "2026-04-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Diallel Analysis

**Complete_diallel_analysis** function Performs Griffing's Half-Diallel Analysis to estimate General Combining Ability (GCA) and Specific Combining Ability (SCA). It fits a mixed model using the \code{sommer}  package to properly pool parental variances, extracts the Best Linear Unbiased  Predictors (BLUPs), and generates a comprehensive visualization using \code{ggplot2}.

```{r getpackage}
library(devtools)
devtools::load_all("/Users/umeshrosyara/Documents/githubdir/DataIntellEngine/r_core/")
testfunction(4,8)
```

```{r analysisflow}
# 1. Load the required libraries
library(sommer)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)

# (Make sure you have already run the complete_diallel_analysis function code here)

# 2. Load the Partial Diallel Data in Wide Format
# This matches the DATALINES from your SAS code
partial_diallel_wide <- read.table(text = "
P1 P2 R1 R2 R3 R4
1 4 128.10 123.84 92.56 115.28
1 5 128.36 119.84 103.24 129.72
1 6 74.40 70.86 60.94 68.00
2 5 91.52 113.96 87.26 106.98
2 6 59.06 65.62 81.62 86.76
2 7 84.16 109.74 102.14 94.52
3 6 109.86 98.16 93.26 102.26
3 7 117.20 100.28 116.16 112.52
3 8 109.68 116.48 123.92 120.86
4 7 53.40 60.86 74.46 69.08
4 8 53.86 48.30 40.64 44.62
5 8 86.62 94.18 90.32 108.16
", header = TRUE)

# 3. Transform to Long Format
# This replaces the SAS DATA step and array loops
partial_diallel_long <- partial_diallel_wide %>%
  pivot_longer(
    cols = starts_with("R"), 
    names_to = "Rep", 
    values_to = "Yield"
  ) %>%
  # Clean up the "Rep" column to just be the number (e.g., "R1" -> 1)
  mutate(Rep = str_remove(Rep, "R"))

# 4. Run the Analysis
cat("Starting Diallel Analysis...\n")
results <- complete_diallel_analysis(
  data = partial_diallel_long, 
  p1_var = "P1", 
  p2_var = "P2", 
  rep_var = "Rep", 
  resp_var = "Yield", 
  trait_name = "Grain Yield"
)

# 5. View the extracted BLUPs
cat("\n--- General Combining Ability (GCA) ---\n")
print(results$GCA_Estimates)

cat("\n--- Specific Combining Ability (SCA) ---\n")
print(results$SCA_Estimates)
```


